---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# note to self 

Render `README.Rmd` regularly to keep `README.md` up-to-date, 
via `devtools::build_readme()` 

# twostage

<!-- badges: start -->
<!-- badges: end -->

The goal of twostage (which may be renamed into item2stage to be more accurate) is to provide a convenient way to fit composite-level models to item-level data. 
For now, this package focuses on methods that do this in a saturated way; that is, the hypothesis that relationships between a formative construct and other variables in the model (whether formative or reflective constructs, or observed variables) flows only through the formative composite is not tested. Furthermore, there is currently now way to adapt these methods to include such tests. That means that these methods are primarily for handling missing data at the item level, when the model is at the composite level; they serve no other purpose and will return identical results to a model that is just fit to the composites instead of the raw items when the data are complete.

The package currently includes the following methods: 
  1) Two-stage ML (TSML) method of Savalei and Rhemtulla (2017a) 
  2) PIM (Pseudo-Indicator Model) method of Rose, Wagner, Mayer, and Nagengast (2019)
  3) (yet to be implemented) the GLS method of Savalei and Rhemtulla (2017b). 
  
  
## To-do list  

Soon: 

1) Provide `show()` and `summary()` for `twostage()` function only (not for component functions), that looks much like lavaan output, but with ML replaced with TSML, test statistic and p-value replaced with res test statistic and p-value, and standard errors replaced with two-stage standard errors (and labeled that way)

2) Add GLS

3) Reconsider how documentation is done for TSML functions -- maybe join, use inherit

4) Add robust TSML (Actually, I think this will just work if we pass `MLR` to lavaan, but check, write a vignette)

5) Reconsider if constructing C is the best way to store info on composites vs components (though it is needed for computations in TSML and GLS, but not in PM). E.g., maybe it should be a list. 

Longer range?:

1) Provide accurate fit indices for model fit assessments that assess the composite-level model only. For the PIM, this may have to be nested fit indices, e.g., $\text{RMSEA}_\text{D}$. How to do fit assessment for these models is probably at least one paper. The fit assessment is a limited of the default PIM implementation, as is noted in the original article. 

2) Implement methods that do test hypotheses about relationships flowing only through composites, i.e., confirmatory composite analysis (CCA), e.g., Schuberth, Henseler, and Ogasawara (2021), Henseler (book, year?). Currently, Yves has provided a preliminary implementation of this approach in lavaan, but it has limitations, such as it cannot handle missing data on items that correspond to exogenous formative latent variables! It also restricts some covariances among items so that constraints imposed by CCA cannot be relaxed to obtain a PIM (these are nested models; see Siegel, Savalei, and Rhemtulla (2025; submitted to PM). 

3) Provide a way for the user to specify how components relate to composites using newly invented lavaan syntax (developed for CCA) for this. Example of such syntax: 

`model <- '
  eta1 <~ x1 + x2
  eta2 <~ x3 + x4 + x5'`
  
This should be translatable to a 5 x 2 matrix C, with rows labeled etas and columns labeled xs, and with 1s corresponding to those xs that belong to certain etas. We could look into how lavaan parses this syntax and stores the result. 


Longer longer longer term goal: 

1) add the two-stage saturated method to lavaan (make it easy for Yves to accept this) and extend current CCA implementation of lavaan so that constraints can be relaxed, yielding PIM


## Installation

You can install the development version of twostage from [GitHub](https://github.com/) with:

``` r
install.packages("pak")
pak::pak("vsavalei/twostage")
```

## Example

This is a basic example:

```{r example}
library(twostage)
## basic example code
```



## Move to a vignette: How to handle observed variables that are not composites

There are two options: 

1) Assign these variables to their own composites, so that single-indicator latent variables are created for them. The advantage of this option is that the generated PIM syntax will be correct right away. The disadvantage is that a different name (from the observed variable's name) has to be assigned to this "composite", making the interpretation of the model results a little awkward. 

2) Include these variables in the model directly as observed variables. The advantage of this option is that the model is more interpretable. The disadvantage is that the generated PIM syntax will have to be modified manually to correlate observed and latent exogenous predictors. This is quite likely since true composites are treated as latent variables in PIM. 

To do so, rename this variable in the composite model, and either regenerate C using stage0 function or rename the corresponding row of C. However, this may require manual modification of the resulting PIM syntax to correlate observed and latent exogeneous predictors (see Github issue #415.

Both approaches are illustrated below. 

```{r}
##PIM syntax tpbmod example

tpbmod<-'
INTALL ~ ATTALL + PBCALL + NORMALL
BEH ~ INTALL'

#made up example with a single-indicator composite
#C1 is the sum of Y1, Y2, and Y3
#' #C2 is the sum of Y4, Y5, and Y6
#' #C3 is Y7
C<-matrix(0,nrow=3,ncol=7)
C[1,1:3]<-1
C[2,4:6]<-1
C[3,7]<-1
rownames(C)<-c("C1","C2","C3")
colnames(C)<-c("Y1","Y2","Y3","Y4","Y5","Y6","Y7")
compmodel<-"C1 ~ C2 + C3"
model1 <- PIM_syntax(C,compmodel)
cat(model1)
data1 <- misdata_mcar20[,c("Y1","Y2","Y3","Y4","Y5","Y6","Y7")]
out <- lavaan::sem(model1,data=data1)
df <- as.numeric(lavaan::fitmeasures(out)["df"])
df

#made up example renaming C3 into Y7:
rownames(C)<-c("C1","C2","Y7")
compmodel<-"C1 ~ C2 + Y7"
model2 <- PIM_syntax(C,compmodel)
cat(model2)
out2 <- lavaan::sem(model2,data=data1) #,orthogonal.x=FALSE) #,fixed.x=FALSE,missing="ML")
df2 <- as.numeric(lavaan::fitmeasures(out2)["df"])
df2
lavaan::lavInspect(out2,"cov.all") #0 for Y7 and C2


model2.mod<-paste(model2,"\n Y7 ~~C2") #Y7~~Y7 doesn't work
out2.mod <- lavaan::sem(model2.mod,data=data1)
out2.mod
#hmm, so why does the default to correlate predictors go away
#when one of the predictors is observed? this is now reported on 
#github, at Yves' request so that he doesn't forget: #414
#as is, we have to explain to the user that this is happening and PIM
#syntax will be incomplete 
```

      
