---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- Render `README.Rmd` regularly to keep `README.md` up-to-date, via `devtools::build_readme()` -->

```{=html}
<!-- To view better, run: 
detach("package:twostage", unload = TRUE)
pkgdown::build_site() 
-->
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Package `twostage`

## Introduction

<!-- badges: start -->

[![R-CMD-check](https://github.com/vsavalei/twostage/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/vsavalei/twostage/actions/workflows/R-CMD-check.yaml)

<!-- badges: end -->

The goal of the `twostage` package is to provide helpful automation to fit composite-level structural equation models (SEMs) to item-level data using the SEM R package `lavaan`. The most common applications are: 1) SEMs with parcels and 2) path analysis with scale scores.

The package includes automation for the following approaches:

1)  The two-stage ML (TSML) method of [Savalei and Rhemtulla (2017a)](https://journals.sagepub.com/doi/full/10.3102/1076998617694880)
2)  The Pseudo-Indicator Model (PIM) of [Rose, Wagner, Mayer, and Nagengast (2019)](https://online.ucpress.edu/collabra/article/5/1/9/112958/Model-Based-Manifest-and-Latent-Composite-Scores)
3)  (yet to be added) the GLS method of [Savalei and Rhemtulla (2017b)](https://www.frontiersin.org/journals/psychology/articles/10.3389/fpsyg.2017.00767/full)

The main purpose of these three approaches is to handle missing data at the item level. When data are complete, these methods/models will return identical (or highly similar, depending on information matrix settings) results compared to the approach where composites are computed directly, and a model is fit to them (see vignette TBA). The composites are always sums or averages of the items; the weights are all equal (for now) and are fixed rather than estimated.

These methods can be contrasted with confirmatory composite analysis (CCA) of [Henseler (2021)](https://www.guilford.com/books/Composite-Based-Structural-Equation-Modeling/J%C3%B6rg-Henseler/9781462545605), which 1) finds optimally weighted composites (for prediction), and 2) does not allow indicators to influence other variables directly (but only through the composite). This approach will *not* return identical results to composite models with complete data. CCA is currently being implemented in `lavaan` (but with some limitations for missing data treatment). 

## Installation

To install twostage from [GitHub](https://github.com/):

``` r
install.packages("pak")
pak::pak("vsavalei/twostage")
```

Load it in the usual way:

```{r}
library(twostage) 
```

## Example

This example uses a simulated dataset `misdata_mcar20`, which contains 27 items, $Y_1$ to $Y_{27}$, where about half have 20% missing data. The model is for composites $C_1$ to $C_9$, which are parcels of three items each, in order; for example, $C_1 = Y_1 + Y_2 + Y_3$, and so on. These composites are never explicitly computed.

The composite model is a 3-factor model, with three indicators each, defined via `lavaan` syntax as follows:

```{r}
#composite model
mod1 <- '
F1 =~ 1*C1 + C2 + C3
F2 =~ 1*C4 + C5 + C6
F3 =~ 1*C7 + C8 + C9
F1 ~~ F1
F2 ~~ F2
F3 ~~ F3
'
```

For now, for true latent variables (*not* composites), the user should set their metric and variances explicitly in the composite model syntax. Above, the first loading of each true latent variable (F1, F2, F3) is set to 1, and their variances are free. This is because the PIM model will be fit via the `lavaan` function, which assumes nothing. The `PIM_syntax` automation function *will* correlate all exogenous variables in the model by default, so it is not necessary to specify, e.g., `F1 ~~ F2` in the composite model syntax (though it doesn't hurt). In general, the user should aim to specify the composite model as explicitly and thoroughly as possible.

To fit this composite model using the item-level methods in this package, the assignment of components to composites needs to be specified. To do this via an interactive interface, use:

```         
C <- stage0(data=misdata_mcar20,model=mod1)
```
This function will first ask the user whether the composites are sums or averages, and then it will ask the user to assign each component to one of the variable names (assumed to be composites) that appear in the model. The following message will confirm the assignment:

```         
Your composites are made up of the following components: 
C1 :  Y1 Y2 Y3 
C2 :  Y4 Y5 Y6 
C3 :  Y7 Y8 Y9 
C4 :  Y10 Y11 Y12 
C5 :  Y13 Y14 Y15 
C6 :  Y16 Y17 Y18 
C7 :  Y19 Y20 Y21 
C8 :  Y22 Y23 Y24 
C9 :  Y25 Y26 Y27 
If this is not correct, start over! 
```
The created matrix $C$ has columns labeled with component names: $Y_1$ to $Y_{27}$, and rows labeled with composite names: $C_1$ to $C_9$. The ijth element of $C$ is nonzero (for sums, it is 1) if component $j$ belongs to composite $i$, and zero otherwise. 

In this file, we create this matrix manually as follows: 

```{r, include = FALSE}
cnames<-lavaan::lavNames(mod1) 
C <- matrix(0,nrow=length(cnames),ncol=length(colnames(misdata_mcar20)))
colnames(C)<-colnames(misdata_mcar20)

rownames(C)<-cnames
C[1,1:3]<-1
C[2,4:6]<-1
C[3,7:9]<-1
C[4,10:12]<-1
C[5,13:15]<-1
C[6,16:18]<-1
C[7,19:21]<-1
C[8,22:24]<-1
C[9,25:27]<-1
```

Once the $C$ matrix is created with the help of the `stage0` function or manually, the composite model can be fit using the methods included in the package.

To fit the composite model using TSML:

```{r}
fit_ts <- twostage(data = misdata_mcar20, compmodel = mod1, C = C)
fit_ts
```

The summary (for now) is custom to `twostage` and does not look like `lavaan`: 
```{r}
summary(fit_ts)
#to see the unadjusted stage 2 summary from lavaan: #selectMethod("summary", "lavaan")(fit_ts)
```

It shows TSML parameter estimates from Stage 2, "naive" standard errors, and TSML standard errors, which are generally larger, reflecting greater uncertainty due to missing data in Stage 1. The (normal theory) residual-based test statistic is also printed. For technical details on the standard errors and the residual-based test statistic computation, see [Savalei and Bentler ( 2009)](https://www.tandfonline.com/doi/full/10.1080/10705510903008238) and [Savalei and Rhemtulla (2017a)](https://journals.sagepub.com/doi/full/10.3102/1076998617694880).

To fit the composite model using PIM, first create the `lavaan` PIM syntax:

```{r}
modpim <- PIM_syntax(compmodel = mod1, C = C)

```

The resulting syntax is long and can be viewed via `cat(modpim)`. It contains the definition of each composite $C_i$, $i=1,\ldots,9$, as a single-indicator latent variable, and a special structure on the components. For details, see [Rose, Wagner, Mayer, and Nagengast (2019).](https://online.ucpress.edu/collabra/article/5/1/9/112958/Model-Based-Manifest-and-Latent-Composite-Scores) It is recommend to always check the part of the generated syntax pertaining to the composites (clearly marked at the bottom). The default in `PIM_syntax` is to covary exogenous variables (whether observed variables, latent variables, or composites); to undo this default, use `exog_cov=FALSE`.

The PIM model can be fit directly in `lavaan`. To treat missing data on the items, use FIML:

```{r}
fit_pim <- lavaan::lavaan(modpim, data=misdata_mcar20,missing="FIML")
fit_pim
```
While the `sem` function would work in this case, it is best to use the `lavaan()` function to fit PIMs to avoid any unintended defaults, given the nonstandard setup of this model. (And until we learn more about it).

Because the PIM setup adds items to the composites model in a saturated way, always confirm that the degrees of freedom are what you would expect for your composite model (i.e., if data had been complete and the composite model had been fit directly). In this example, the composite model is a 3-factor CFA with nine indicators, so it should have $9(10)/2 - (9+9+3)=`r 9*(10)/2 - (6+9+3+3)`$ degrees of freedom, where the parameters are free factor loadings, residual variances, and factor variances and covariances. When fit correctly,`fit_pm` should also return the same df as `fit_ts`. 

To inspect parameter estimates, use `summary(fit_pim)` or the `parameterEstimates` function of `lavaan`. As the interest is in the composite model, the user has to ignore a lot of extraneous output pertaining to the item parameters (here, all rows referring to "Y" variables).

To compare estimates of common parameters in TS and PIM, which are those in the composite model only, we can use the `compare_est` function:

```{r}
comp_table <- compare_est(fit_pim,fit_ts)
comp_table
```

The TS and PIM estimates and standard errors are quite similar. Technically, the TS estimates are less efficient than FIML (PIM) estimates. When data are complete, both TS and PIM produce equivalent output to the complete data run on the manually-formed composites (see the Complete_data vignette).

Approximate fit assessment of PIM models requires special setup, as discussed in [Rose, Wagner, Mayer, and Nagengast (2019).](https://online.ucpress.edu/collabra/article/5/1/9/112958/Model-Based-Manifest-and-Latent-Composite-Scores). The gist of the issue is that we want to assess the fit of the composite model, not the fit of the entire model with items added in a saturated way; this requires recomputing CFI and SRMR (but RMSEA is fine). To get corrected fit measures, use: 

```{r allfit}
fitm_pim <- fitMeasures_pim(C, compmodel=mod1, fit_pim=fit_pim, 
                data = misdata_mcar20)
```
We can compare the PIM and TS fit indices:

```{r allfitcomp}
indices<-c("rmsea","rmsea.robust","cfi","cfi.robust","tli","tli.robust","srmr")
fitm_pim[indices]

fitMeasures(fit_ts,indices)
```

When `estimator ="FIML"`, as in PIM, corrections are necessary to fit indices in order for them to mimic their would-be complete data values [Zhang and Savalei (2023)](https://psycnet.apa.org/doiLanding?doi=10.1037%2Fmet0000445). Thus, we compare the RMSEA from `fit_ts` to `rmsea.robust` from `fit_pim`. Similarly, we should compare the CFI from `fit_ts` to the `cfi_robust` in ` fit_pim`, but this CFI is custom-computed using a different baseline model, so it is not clear if these corrections are right. No corrections are necessary to SRMR, so these can be compared directly. See the [Approximate_fit vignette](../doc/Approximate_fit.html) for more detail. Lastly, TSML indices require small-sample corrections (these are less consequential than PIM corrections, as they go away asymptotically), but these have not yet been implemented. 

```{=html}
<!-- Links to (../doc/Complete_data.html) do not work until the package website is on Github. Links to  (../vignettes/Complete_data.html) do not work because this folder does not contain hmtl files. Ignore for now, see advice below.

<!-- Note on GitHub: If the README is on GitHub and you want to link to the rendered vignette during development, you could manually provide a link to a rendered version stored externally (e.g., on GitHub Pages) until it is published on CRAN.-->
```
