---
title: "Approximate Fit Assessment with Composite Models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Approximate_fit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
#Use devtools::build_vignettes() to build the vignettes. 
```

In this vignette we assume that the interest is in the composite-level model; that is, PIM is used to treat missing data, and the covariance structure of the items is not of interest. In this case, adjustments are required to some fit indices to remove the influence of the raw items on fit, as discussed in [Rose, Wagner, Mayer, and Nagengast (2019).](https://online.ucpress.edu/collabra/article/5/1/9/112958/Model-Based-Manifest-and-Latent-Composite-Scores) This vignette provides more detail on the implemented adjustments.

The goal of the adjustments is to estimate what the fit would have been had there been no missing data. By default, incremental fit indices (such as CFI and TLI) obtained by applying `fitMeasures()` function to a fitted PIM will be evaluating the fit of PIM against the traditional baseline or null model, which is the independence model that un-correlates all items and removes the latent variables (i.e., the composites) from the model entirely. Instead, to appropriately assess the incremental fit of the composite-level model, we need to set up a different baseline model -- one in which the items are freely correlated, but the composites, which remain in the model set up as latent variables with a PIM structure on the items, are not. Similarly, the default SRMR will compare observed correlations among the *items* to model-implied correlations among the items. Instead, to evaluate the fit of the composite-level model only, the adjusted SRMR needs to capture the residuals in estimated *composite* relationships between the structured (H0) and the saturated (H1) model for *composites*. This requires a special H1 model setup.

## Example 1. Complete Data

We first illustrate the composite-level fit index adjustments on a complete data example (from the complete data vignette). If our adjustments to CFI/TLI and SRMR are correct, the adjusted versions of these indices should match what we get when we just fit the composite-level model to the composites computed from complete data:

```{r compdata}
library(twostage)

tpbdatac<-data.frame(matrix(ncol = 0, nrow = nrow(tpbdata)))
at_names <- grep("AT", names(tpbdata), value = TRUE) 
tpbdatac$ATTALL<-rowSums(tpbdata[at_names]) 
tpbdatac$PBCALL<-rowSums(tpbdata[c("PBC1","PBC2","PBC3")]) 
tpbdatac$NORSALL<-rowSums(tpbdata[c("NORS1","NORS2","NORS3")]) 
tpbdatac$INTALL<-rowSums(tpbdata[c("INT1","INT2","INT3")]) 
tpbdatac$BEH<-rowSums(tpbdata["BEH"]) 


#composite-level model (BEH stays as as single indicator variable)
tpbmod<-'
INTALL ~ ATTALL + PBCALL + NORSALL
BEH ~ INTALL
'

fit_comp<-lavaan::sem(tpbmod,data=tpbdatac,fixed.x=FALSE,meanstructure=TRUE)
fit_comp_base<- lavaan:::lav_object_independence(fit_comp)
fitm_comp <- lavaan::fitMeasures(fit_comp)[c("npar","chisq","df","baseline.chisq",
                                   "baseline.df","cfi","tli","rmsea","srmr_mplus",
                                   "srmr_mplus_nomean")]
round(fitm_comp,3)
```

We should obtain fit indices that are identical to these values. The chi-square difference test between the proposed and the baseline model should also be replicated.

Below we construct the C matrix, create PIM syntax, and fit PIM to the raw items in `tpbdata` (without declaring missing data in this case, as there is none):

```{r}
cnames<-lavaan::lavNames(tpbmod)
C <- matrix(0,nrow=length(cnames),ncol=length(colnames(tpbdata)))
colnames(C)<-colnames(tpbdata)
rownames(C)<-cnames
C[1,c("INT1","INT2","INT3")]<-1
C[2,c("BEH")]<-1
C[3,grep("AT", names(tpbdata))]<-1
C[4,c("PBC1","PBC2","PBC3")]<-1
C[5,c("NORS1","NORS2","NORS3")]<-1

tpbpim <- PIM_syntax(compmodel=tpbmod,C=C)
fit_pim <- lavaan::lavaan(tpbpim, data=tpbdata)
fitm_pim <- lavaan::fitMeasures(fit_pim)[c("npar","chisq","df","baseline.chisq","baseline.df","cfi","tli","rmsea","srmr_mplus", "srmr_mplus_nomean")]
round(fitm_pim,3)
```

While the model chi-square, model df, and the RMSEA agree between `fit_pim` and `fit_comp`, the baseline model fit, the CFI/TLI, and SRMR differ.

#### Getting the composite-level incremental fit indices

To get the appropriate composite-level CFI/TLI, we set up a new baseline model that only un-correlates variables in the composite model, while leaving the rest of the PIM structure intact. The function `PIM_syntax_base` will create such a PIM model:

```{r base1}
tpbpim_base <- PIM_syntax_base(compmodel=tpbmod, C=C, exog_cov=TRUE)
#cat(tpbpim_base)
```

The user should verify the model is set up as intended by running the `cat()` command above. While the PIM syntax is long, the composite model part is clearly flagged with:

```         
##--------Composite Model (inspect carefully): ----------##
```

By default, the three exogenous predictors remain correlated, which can be undone by specifying `exog_cov=FALSE`. Correlated exogenous predictors are usually the default in `lavaan` if the exogenous variables have not been mentioned in the syntax except as predictors.\footnote{In `lavaan`, the default baseline model will change if any of the exogenous variables are mentioned in the syntax by specifying their variance or a covariance; in this case, their correlations in the baseline model will be set to zero instead of freely estimated. This default can be confusing, but it is `lavaan`'s attempt to keep the baseline model nested within the fitted model.}

We then fit the PIM baseline model and verify that its chi-square and df match the composite-level baseline model for complete data:

```{r base2}
fit_pim_base <- lavaan::lavaan(tpbpim_base, data=tpbdata)

lavaan::lavTestLRT(fit_pim,fit_pim_base)
lavaan::lavTestLRT(fit_comp,fit_comp_base)
```

To compute CFI/TLI based on this baseline model, we can provide this custom baseline model to the `fitMeasures` function in `lavaan`:

```{r base3}
pim.incremental <- lavaan::fitMeasures(fit_pim, baseline.model=fit_pim_base,fit.measures=c("cfi","tli"))
pim.incremental

#comparison to complete data:
round(fitm_comp[c("cfi","tli")],3)

```

These CFI and TLI are identical to the values for complete data.

#### Getting the composite-level SRMR

The correct SRMR should capture the difference in the model-reproduced covariances between the saturated (H1) and structured (H0) model for the composites To get the correct SRMR for the composite-level model, we first need to fit a PIM H1, which will freely inter-correlate the variables in the composites model, while keeping the PIM structure on the items. The function `PIM_syntax_sat` will take H0 composite-level model as input (`tpbmod`), turn it into a saturated model, and combine with the PIM structure on the items:

```{r srmr1}
tpbpim_sat <- PIM_syntax_sat(C=C,compmodel=tpbmod)
fit_pim_sat <- lavaan::lavaan(tpbpim_sat, data=tpbdata)

```

The resulting model should have zero degrees of freedom.

The correct SRMR should be based on the differences between the model-implied covariances, for the composite-level model variables only, between `fit_pim` and `fit_pim_sat`, which are shown below:

```{r srmr2}
comp_names <- lavaan::lavNames(tpbmod)

Sigma_h1 <- lavaan::lavInspect(fit_pim_sat,"cov.all")[comp_names, comp_names]
Sigma <- lavaan::lavInspect(fit_pim,"cov.all")[comp_names, comp_names]
Res <- Sigma_h1 - Sigma
round(Res,3)
```

We compute the correct composite-level SRMR using the function below (which is adapted from lavaan's function `lav_fit_srmr_mplus`):

```{r srmr3}
srmr <- srmr_mplus_pim(fit_pim,fit_pim_sat,tpbmod)
srmr

#comparison to complete data:
round(fitm_comp[c("srmr_mplus","srmr_mplus_nomean")],3)
```

Both SRMR without the means and one with the means match the corresponding values from complete data. Because PIM models were developed for missing data, `meanstructure=TRUE` is likely always on when these models are used; however, if a model has a saturated mean structure (as is the case here), it is better to use the no-mean version of SRMR that does not average over mean structure residuals (which will be zero or very close to it, artificially improving the SRMR).

#### Obtaining all composite-level indices at once

The above discussion was mostly for instructional purposes. The following function will obtain all fit measures for PIM at once, as well as print those that are correct without an adjustment (such as the RMSEA). 

```{r allfit}
fitm_pim <- fitMeasures_pim(C, compmodel=tpbmod, fit_pim=fit_pim, exog_cov_base = TRUE)
fitm_pim[c("chisq","df","baseline.chisq","baseline.df","cfi","tli","rmsea","srmr_mplus", "srmr_mplus_nomean")]

#compare to complete data
fitm_comp[c("chisq","df","baseline.chisq","baseline.df","cfi","tli","rmsea","srmr_mplus", "srmr_mplus_nomean")]
```

Better function with fewer arguments will be written later.

## Example 2. Incomplete Data

With missing data, FIML fit indices CFI/TLI and RMSEA underestimate the amount of model misfit, relative to what it would have been had the data been complete [Zhang and Savalei (2020)](https://www.tandfonline.com/doi/full/10.1080/10705511.2019.1642111). As PIM is a model that uses FIML in a clever way to treat item-level missing data, the PIM RMSEA and CFI/TLI (computed using the corrected baseline model, as shown above) will be biased downward. However, *lavaan* automatically adjusts for this bias and prints robust versions of CFI/TLI/RMSEA with FIML, following [Zhang and Savalei (2023)](https://psycnet.apa.org/record/2022-18510-001).

To illustrate the impact of these corrections, we will use the example from the README file, but misspecify the composite model. The composite model is a 3-factor model with 9 composite indicators, but we will force one of the factor correlations to be zero. The data are on 27 items. Below we set up the composite model and the C matrix, and fit the composite model using PIM:

```{r}
mod1a <- '
F1 =~ 1*C1 + C2 + C3
F2 =~ 1*C4 + C5 + C6
F3 =~ 1*C7 + C8 + C9
F1 ~~ F1  
F2 ~~ F2
F3 ~~ F3
F1 ~~ 0*F2 #added to make the model false

'
cnames<-lavaan::lavNames(mod1a) 
C <- matrix(0,nrow=length(cnames),ncol=length(colnames(misdata_mcar20)))
colnames(C)<-colnames(misdata_mcar20)

rownames(C)<-cnames
C[1,1:3]<-1
C[2,4:6]<-1
C[3,7:9]<-1
C[4,10:12]<-1
C[5,13:15]<-1
C[6,16:18]<-1
C[7,19:21]<-1
C[8,22:24]<-1
C[9,25:27]<-1

modpim <- PIM_syntax(compmodel = mod1a, C = C)
fit_pim <- lavaan::lavaan(modpim, data=misdata_mcar20,missing="FIML")
```

To use a custom baseline model in the computation of the CFI/TLI:

```{r}
mod_base <- PIM_syntax_base(C=C,compmodel=mod1a)
#cat(mod_base) #always check! 
fit_pim_base <- lavaan::lavaan(mod_base, data=misdata_mcar20,missing="FIML")

fitm_pim_incr <- lavaan::fitMeasures(fit_pim, 
                                       baseline.model=fit_pim_base,   fit.measures=c("rmsea","rmsea.robust","cfi", "tli","cfi.robust","tli.robust"))
fitm_pim_incr 
```

In the output above, the robust RMSEA, CFI, and TLI have FIML missing data adjustments incorporate (so they are "robust" to missing data). They reflect poorer fit relative to the uncorrected versions, because missing data tends to artificially inflate model fit.

While SRMR does not require missing data adjustments, for completeness, let's get the correct composite-level SRMR.

```{r}
pim_sat <- PIM_syntax_sat(compmodel=mod1a,C=C)
fit_pim_sat <- lavaan::lavaan(pim_sat, data=misdata_mcar20,missing="FIML")

srmr <- srmr_mplus_pim(fit_pim,fit_pim_sat,mod1a)
srmr
```

# Obtaining Corrected TSML Fit Indices

## Example 1. Complete Data

As illustrated in the [Complete_data vignette](../doc/Complete_data.html), with complete data, fit measures for the direct composites run and the naive Stage 2 fit measures for the TSML run are the same. Thus, there is nothing to illustrate!

## Example 2. Incomplete Data

```{r}
fit_ts <- twostage(data = misdata_mcar20, compmodel = mod1a, C = C)
#fitMeasures(fit_ts)
```

When it comes to TSML-based RMSEA and CFI/TLI, unlike their FIML counterparts, their asymptotic values are already correct. This is because with incomplete data, the "naive" TSML estimates of RMSEA, CFI, etc will be based on the fit function value from Stage 2, which is already the complete data fit function.

However, corrections are still needed to improve their small-sample performance, as discussed in [Zhang and Savalei (2023)](https://psycnet.apa.org/doiLanding?doi=10.1037%2Fmet0000445), though this article did not consider item-level TSML. These small sample adjustments basically amount to subtracting more than just df to get the non-centrality parameters in RMSEA/CFI/TLI computations. (See Appendix). These adjustments are automatically done in the current version of `twostage`, and are reflected in the robust versions of RMSEA, CFI, and TLI printed by `fitMeasures`.

The correct robust fit measures for TS and PIM are quite similar:

```{r}
fitMeasures(fit_ts)[c("rmsea.robust","cfi.robust","tli.robust","srmr")]

fitm_pim <- fitMeasures_pim(C=C, compmodel = mod1a, fit_pim=fit_pim) #there are no exogeneous C's so this makes no difference

fitm_pim[c("rmsea.robust","cfi.robust","tli.robust","srmr")]
```

# Appendix: Technical Details for Small Sample Corrections to TSML Fit Indices

A note on notation: In Savalei & Rhemtulla (2017), we used the subscript $c$ to refer to composite variables. In Zhang & Savalei (2020; 2023), we used it to refer to complete data quantities. Here, I will use subscript $p$ to refer to quantities based on the items, and no subscript for quantities based on composites (that is, I'm dropping subscript $c$ we used in 2017). For fit functions, I'm keeping the notation of Zhang & Savalei (2023), where $c$ refers to complete-data fit function and various matrices computed from it. Because complete data tends to go with composites, hopefully this is not too confusing.

Assume we have $p$ items and $k$ composites. In Stage 1 of TSML, we use FIML to obtain $\tilde\mu_p$ ($p \times 1$) and covariance matrix $\tilde\Sigma_p$ ($p \times p$), the saturated "EM" estimates of item means and covariances. We also obtain the asymptotic covariance matrix of these means and covariances, $\tilde\Gamma_p$ ($p^* \times p^*)$, where $p* = p(p+1)/2+p$. Under normality, we can estimate $\tilde\Gamma_{p,N}=\tilde A_p^{-1}$, where $\tilde A_p$ is observed Fisher information for the saturated model for the items. We can also obtain a more general estimate $\tilde\Gamma_p=\tilde W^{-1}_p \tilde V_p \tilde W^{-1}_p$, valid for nonnormal data, where $\tilde V_p$ is the first-order information matrix (see Yuan & Bentler, 2000; Savalei & Rosseel, 2021).

In Stage 1a of TSML, these quantities are reduced to the $k$ dimensions using a selection matrix $C$. The saturated estimates of means and variances of the composites are: $\tilde\mu=C \tilde\mu_p$, $\tilde\Sigma=C \tilde\Sigma_p C'$, and their asymptotic covariance matrix is \$ \tilde\Gamma= C\_{\text{big}} \tilde\Gamma\*p C'\*{\text{big}}\$, respectively, where $C_{\text{big}}$ is constructed from $C$ (see Savalei & Rhemtulla, 2017, for details).

In Stage 2, TS parameter estimates are obtained by minimizing $F_{\text{c}}(\Sigma,\mu|\tilde\Sigma,\tilde\mu)$, the complete-data ML fit function with saturated FIML estimates of means and covariances of composites from Stage 1a as input. Denote the resulting TSML estimates by $\breve\Sigma=\Sigma(\breve\theta)$ and $\breve\mu=\mu(\breve\theta)$, and the resulting fit function minimum by $\breve  F_{\text{c}}$. For the baseline model needed for the CFI, denote the resulting TS model-reproduced estimates by $\breve\Sigma_B$ and $\breve\mu_B$, and the resulting fit function minimum by $\breve  F_{\text{c,B}}$. Below $n$ refers to total sample size, rather than $n-1$, to be consistent with default behavior of `lavaan` (this represents a minor difference from prior published papers).

The corrected TSML fit indices are computed as follows: $$\text{RMSEA}_{\text{TS}} = \sqrt{\frac{ \breve  F_{\text{c}}}{df}-\frac{c}{df \times n}}$$ $$\text{CFI}_{\text{TS}}=1-\frac{\breve  F_{\text{c}} -\frac{c}{n}}{\breve  F_{\text{c},B}-\frac{c_B}{n}},$$ where $c=tr[\tilde U_\text{c} \tilde\Gamma]$, $c_B=tr[\tilde U_\text{c,B}]$, $\tilde\Gamma$ is the asymptotic covariance matrix that comes from Stage 1a (under normality, it is replaced with $\tilde\Gamma_N$), $\tilde U_c = \tilde  W_c - \tilde W_c \breve \Delta (\breve \Delta' \tilde W_c \breve \Delta)^{-1} \breve \Delta' \tilde W_c$, $\tilde U_{c, B} = \tilde W_{c} - \tilde W_{c} \breve \Delta_B (\breve \Delta_B' \tilde W_{c} \breve \Delta_B)^{-1} \breve \Delta_B' \tilde W_{c}$, $\tilde W_c$ is the "naive" complete-data information matrix for the saturated model obtained in Stage 2, evaluated at the saturated model estimates, and $\breve\Delta$ and $\breve \Delta_B$ are the matrices of model derivatives for hypothesized (H0) and baseline model, respectively, evaluated at the TSML estimates.

These equations are essentially the same as in Zhang & Savalei (2023), because the generalization to item-level TSML only requires that $\tilde \Gamma$ is based on composites and not the raw items.

Other computational versions of the corrections $c$ and $c_B$, such as those based on structured estimates (Savalei & Rosseel, 2022), are possible. Zhang & Savalei (2023), who studied nonnormal versions of TSML corrections, found that the version based on saturated estimates performed the best (TS2, in their study).

<!-- What is implemented in `twostage` via `lavaan`?-->
