---
title: "Handing Observed Variables that are not Composites in PIM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Handing_observed_vars_that_arent_composites}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(twostage)
packageVersion("lavaan")
```

## How to handle observed variables that are not composites in PIM

There are two options: 

1) Assign these variables to their own composites, so that single-indicator latent variables are created for them. A different name (from the observed variable's name) has to be assigned to this composite. This option is included for completeness. 

2) Include these variables in the model directly as observed variables. 

Both approaches are illustrated below on two examples: one with an endogenous and one with an exogenous observed variables.  

(This vignette is mostly to ensure PIM sets up correct syntax in unusual cases.)

## Example 1: TPB Data

The `tpbmod` model fit to `tpbdata` (see Complete Data vignette) has `BEH` as an endogenous observed variable:  

```{r}
library(twostage)

tpbmod<-'
INTALL ~ ATTALL + PBCALL + NORMALL
BEH ~ INTALL'
```

The complete data vignette illustrated the second approach by creating the following matrix C:

```{r C}
cnames<-lavaan::lavNames(tpbmod)
C <- matrix(0,nrow=length(cnames),ncol=length(colnames(tpbdata)))
colnames(C)<-colnames(tpbdata)
rownames(C)<-cnames
C[1,c("INT1","INT2","INT3")]<-1
C[2,c("BEH")]<-1
C[3,grep("AT", names(tpbdata))]<-1
C[4,c("PBC1","PBC2","PBC3")]<-1
C[5,c("NORS1","NORS2","NORS3")]<-1
C
```
Here, `BEH` is the name of observed variable, but it is also the name of one of the composites. Because a row in C has the same name as a column of C, this variable will be treated as an observed variable and no latent variable will be created for it. A message is printed to that effect: 
```{r}
tpbpim <- PIM_syntax(compmodel=tpbmod,C=C)
```

The overall model syntax is long but we can pull out lines pertaining to the `BEH` variable to verify: 
```{r}
#cat(tpbpim) #the whole syntax, long
# Extract lines containing "BEH"
syntax_lines <- strsplit(tpbpim, "\n")[[1]]
beh_lines <- grep("BEH", syntax_lines, value = TRUE)
cat(paste(beh_lines, collapse = "\n"))

```
The part of the syntax where `BEH` is correlated with all (but one for each composite) indicators of the other composites is necessary for PIM.

Alternatively, we can create a single-component composite to represent BEH. For this, we would need to name this composite something other than the name of the observed variable: 

```{r}
#use a different composite name for BEH
tpbmod1<-'
INTALL ~ ATTALL + PBCALL + NORMALL
BEH_lat ~ INTALL'

C1 <- C
cnames1<-lavaan::lavNames(tpbmod1)
rownames(C1)<-cnames1
C1

tpbpim1 <- PIM_syntax(compmodel=tpbmod1,C=C1)

# Extract lines containing "BEH"
syntax_lines1 <- strsplit(tpbpim1, "\n")[[1]]
beh_lines1 <- grep("BEH", syntax_lines1, value = TRUE)
cat(paste(beh_lines1, collapse = "\n"))
```
We now have a single indicator latent variable `BEH_lat` corresponding to `BEH`. 

## Example 2

This is a small artificial example using variables in `misdata_mcar20`. The model makes no sense; we are just testing that exogenous composites are treated correctly no matter how they are specified. There is one exogenous observed variable, $Y_1$ (or $C_1$). There is also an exogenous composite, $C_2$.  Together, they are predicting a composite $C_3$.  

PIM syntax when treating $Y_1$ as an observed variable: 
```{r}
#made up example with a single-indicator composite
#C1 is Y1
#C2 is the sum of Y2, Y3, and Y4
#C3 is Y5 and Y6
C<-matrix(0,nrow=3,ncol=6)
C[1,1]<-1
C[2,2:4]<-1
C[3,5:6]<-1

#as an observed variable
rownames(C)<-c("Y1","C2","C3")
colnames(C)<-c("Y1","Y2","Y3","Y4","Y5","Y6")
compmodel<-"C3 ~ Y1 + C2"
model <- PIM_syntax(compmodel = compmodel, C = C)
cat(model)
```
  

PIM syntax when treating $Y_1$ as a single indicator composite $C_1$: 
```{r}
C1<-matrix(0,nrow=3,ncol=6)
C1[1,1]<-1
C1[2,2:4]<-1
C1[3,5:6]<-1
rownames(C1)<-c("C1","C2","C3")
colnames(C1)<-c("Y1","Y2","Y3","Y4","Y5","Y6")
compmodel1<-"C3 ~ C1 + C2"
model1 <- PIM_syntax(compmodel = compmodel1, C = C1)
cat(model1)
```

We can fit both these models to data with the same results: 
```{r}
fit <- lavaan::lavaan(model, data=misdata_mcar20,missing="FIML")
fit1 <- lavaan::lavaan(model1, data=misdata_mcar20,missing=
                         "fiml")
compare_est(fit,fit1) #commonly labeled parameters

```

To both illustrate the use of `exog_cov` option and to test that the fit statistics are also the same, let's fit a model that is not saturated by forcing predictors to correlate at zero: 

```{r}
model <- PIM_syntax(compmodel = compmodel, C = C1,exog_cov=FALSE)
model1 <- PIM_syntax(compmodel = compmodel1, C = C1,exog_cov=FALSE)

fit <- lavaan::lavaan(model, data=misdata_mcar20,missing="FIML")
fit1 <- lavaan::lavaan(model1, data=misdata_mcar20,missing=
                         "fiml")

indices<-fitMeasures_pim(C,compmodel,fit,exog_cov_base=FALSE) 

indices1<-fitMeasures_pim(C1,compmodel1,fit1,exog_cov_base=FALSE) 

rbind(indices,indices1)

#compare_est(fit,fit1) #commonly labeled parameters

```

