---
title: "Handing_observed_vars_that_arent_composites"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Handing_observed_vars_that_arent_composites}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(twostage)

```
<!--#Note to self: use devtools::build_rmd("vignettes/my-vignette.Rmd") to render the #vignette. This builds your vignette against a (temporarily installed) development #version of your package. -->


## Move to a vignette: How to handle observed variables that are not composites

There are two options: 

1) Assign these variables to their own composites, so that single-indicator latent variables are created for them. The advantage of this option is that the generated PIM syntax will be correct right away. The disadvantage is that a different name (from the observed variable's name) has to be assigned to this "composite", making the interpretation of the model results a little awkward. 

2) Include these variables in the model directly as observed variables. The advantage of this option is that the model is more interpretable. The disadvantage is that the generated PIM syntax will have to be modified manually to correlate observed and latent exogenous predictors. This is quite likely since true composites are treated as latent variables in PIM. 

To do so, rename this variable in the composite model, and either regenerate C using stage0 function or rename the corresponding row of C. However, this may require manual modification of the resulting PIM syntax to correlate observed and latent exogenous predictors (see Github issue #415.

Both approaches are illustrated below. 

```{r}
##PIM syntax tpbmod example

tpbmod<-'
INTALL ~ ATTALL + PBCALL + NORMALL
BEH ~ INTALL'

#made up example with a single-indicator composite
#C1 is the sum of Y1, Y2, and Y3
#' #C2 is the sum of Y4, Y5, and Y6
#' #C3 is Y7
C<-matrix(0,nrow=3,ncol=7)
C[1,1:3]<-1
C[2,4:6]<-1
C[3,7]<-1
rownames(C)<-c("C1","C2","C3")
colnames(C)<-c("Y1","Y2","Y3","Y4","Y5","Y6","Y7")
compmodel<-"C1 ~ C2 + C3"
model1 <- PIM_syntax(C,compmodel)
cat(model1)
data1 <- misdata_mcar20[,c("Y1","Y2","Y3","Y4","Y5","Y6","Y7")]
out <- lavaan::sem(model1,data=data1)
df <- as.numeric(lavaan::fitmeasures(out)["df"])
df

#made up example renaming C3 into Y7:
rownames(C)<-c("C1","C2","Y7")
compmodel<-"C1 ~ C2 + Y7"
model2 <- PIM_syntax(C,compmodel)
cat(model2)
out2 <- lavaan::sem(model2,data=data1) #,orthogonal.x=FALSE) #,fixed.x=FALSE,missing="ML")
df2 <- as.numeric(lavaan::fitmeasures(out2)["df"])
df2
lavaan::lavInspect(out2,"cov.all") #0 for Y7 and C2


model2.mod<-paste(model2,"\n Y7 ~~C2") #Y7~~Y7 doesn't work
out2.mod <- lavaan::sem(model2.mod,data=data1)
out2.mod
#hmm, so why does the default to correlate predictors go away
#when one of the predictors is observed? this is now reported on 
#github, at Yves' request so that he doesn't forget: #414
#as is, we have to explain to the user that this is happening and PIM
#syntax will be incomplete 

# title: this is the title that appears in the vignette. If you change it, make sure to make the same change to VignetteIndexEntry{}. They should be the same, but unfortunately that’s not automatic.

```

<!-- Note to the self: I just realized that this function has no option for excluding
observed variables that are not composites, if they appear in the composites model. 
However, if they are observed variables, they will also be in the dataset, so they can
be assigned to their own `composite`. But the user may feed only a subportion of the data
that truly are components: we need to have informative messages to guide all this.-->   
